Library Learning for Neurally-Guided Bayesian Program Induction

### 2018-10-31 nips_2018.tex
version space archive progress

\title{Library Learning in Neurally-Guided Bayesian Program Induction}

@@

\textbf{Compile: learning to tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.
We train $q$ both on programs found during the Explore step and
on samples from the current DSL.

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
  \end{scope}}
    \end{tikzpicture}
  \caption{How these steps bootstrap each other.}  \label{feeding}\vspace{-1cm}
\end{wrapfigure}%
Crucially, each of these three steps bootstraps the others (Fig.~\ref{feeding}): improving either the DSL or the recognition model
makes search easier, so we find more programs solving tasks;
both improving the DSL and solving more tasks expands the training data for the recognition model;
and finding more programs that solve tasks
gives more data from which to learn a DSL.

@@

%% Learning the DSL eases the difficulty of synthesis
%% by exposing a domain-specific basis for constructing programs.
%% Another complementary means of meeting search is to learn a
%% bottom-up \emph{recognition model}
\vspace{0.5cm}Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
For added robustness,
we interleave an extra program search step (Explore) before training the recognition model,
and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.

@@

@@

\begin{wrapfigure}{r}{0.45\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-2cm}
\end{wrapfigure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

@@

\begin{figure}\centering
  \includegraphics[width = 4.35cm]{figures/listLearningCurve_color.eps}
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves}
\end{figure}

@@

 \noindent \textbf{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, Bayesian program merging~\cite{hwang2011inducing}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
 Concurrent with this work, closely related ideas were applied to mining `code idioms' from imperative programs~\cite{8355713}
 and for synthesizing functional programs from natural language~\cite{Alex}.

@@

\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, the MUSE program (Darpa grant FA8750-14-2-0242),
and an AWS ML Research Award.

>>>>

\title{Library Learning for Neurally-Guided Bayesian Program Induction}

@@

\textbf{Compile: learning to tractably maximize $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.
We train $q$ both on programs found during the Explore step and
on samples from the current DSL.

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,inner sep=0pt,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Explore)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,inner sep=0pt,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Explore)}} (synthesis.east);
  \end{scope}}
    \end{tikzpicture}
  \caption{How these steps bootstrap each other.}  \label{feeding}\vspace{-1cm}
\end{wrapfigure}%
Crucially, each of these three steps bootstraps the others (Fig.~\ref{feeding}): improving either the DSL or the recognition model
makes search easier, so we find more programs solving tasks;
both improving the DSL and solving more tasks expands the training data for the recognition model;
and finding more programs that solve tasks
gives more data from which to learn a DSL.

@@

%% Learning the DSL eases the difficulty of synthesis
%% by exposing a domain-specific basis for constructing programs.
%% Another complementary means of meeting search is to learn a
%% bottom-up \emph{recognition model}

@@

%\vspace{0.5cm}
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
For added robustness,
we interleave an extra program search step (Explore) before training the recognition model,
and just enumerate from the prior $(\mathcal{D},\theta)$ during this extra Explore step.

@@

\begin{wrapfigure}{r}{0.55\textwidth}\vspace{-0.5cm} \newcommand{\functionSize}{1.5cm}\centering
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/48.png}
  \includegraphics[width = \functionSize]{figures/functions/102.png}
  \includegraphics[width = \functionSize]{figures/functions/116.png}\\
  \vspace{1pt}\includegraphics[width = \functionSize]{figures/functions/181.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/149.png}
    \includegraphics[width = \functionSize]{figures/functions/187.png}
  \caption{Recognition model input for symbolic regression. DSL learns subroutines for polynomials (top row) and rational functions (bottom row) while the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines best explains the observation.}\label{functions}\vspace{-0.5cm}
\end{wrapfigure}
\system learns a DSL containing 13 new functions,
most of which are templates for polynomials of different orders or ratios of polynomials.
It also learns to find programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_4$ in the rightmost column of Fig.~\ref{initialExampleDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing --- both the implicit bias towards shorter programs and the likelihood model's BIC penalty.

@@

\begin{figure}[h]\centering
  \includegraphics[width = 4.35cm]{figures/listLearningCurve_color.eps}
  \includegraphics[width = 4.5cm]{figures/textLearningCurve.eps}
  \includegraphics[width = 4.5cm]{figures/rationalCurve.eps}
  \caption{Learning curves for \system both with (\orange{in orange}) and without
  (\teal{in teal}) the recognition model. Solid lines: \% holdout testing tasks solved. Dashed lines: Average solve time.}\label{learningCurves}
\end{figure}

@@

 \noindent \textbf{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10}, Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}, MagicHaskeller~\cite{katayama2015towards}, Bayesian program merging~\cite{hwang2011inducing}, and predicate invention in Inductive Logic Programming~\cite{DBLP:conf/ecai/LinDETM14}.
 Closely allied ideas have been applied to mining `code idioms' from  programs~\cite{Allamanis:2014:MIS:2635868.2635901}, and, concurrent with this work, using those idioms to better synthesize functional programs from natural language~\cite{Alex}.

@@

\subsubsection*{Acknowledgments} We are grateful for collaborations with Eyal Dechter, whose EC algorithm directly inspired this work, and for funding from the NSF GRFP, AFOSR award FA9550-16-1-0012, the MIT-IBM Watson AI Lab, the MUSE program (Darpa grant FA8750-14-2-0242),
and an AWS ML Research Award.
This material is based upon work supported by the Center for Brains, Minds and Machines (CBMM), funded by NSF STC award CCF-1231216.
